---
title: "Class 09: Halloween"
author: "Destiny (A16340362)"
format: pdf
---


```{r}
candy_file <- "candy-data (1).csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```
>Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

>Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Starburst", ]$winpercent
```

>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent

```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
library(skimr)
skim(candy)
```
>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

-Yes there are variables/ columns that do look like they're on a different scale. Fpr example for the variable winpercent 

>Q7. What do you think a zero and one represent for the candy$chocolate column?

-The one represents that the candy does contain chocolate, but the zero means that the candy doesn't contain chocolate

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

>Q9. Is the distribution of winpercent values symmetrical?

-No, the winpercent values aren't symmetrical 

>Q10. Is the center of the distribution above or below 50%?

-The distribution of the center is below 50%

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

-On average, chocolate is ranked higher than the fruit candy.

```{r}
candy$chocolate
choc.Inds <- as.logical(candy$chocolate)
choc.win <- candy[choc.Inds, ]$winpercent
mean(choc.win)
```

```{r}
fruity.Ind <- as.logical(candy$fruity)
fruity.win <- candy[fruity.Ind, "winpercent"]
mean(fruity.win)
```

>Q12. Is this difference statistically significant?

-The difference is statistically significant 

```{r}
t.test(choc.win, fruity.win)
```
>Q13. What are the five least liked candy types in this set?

-Nik L Nip, Boston Baked Beans, Chiclets, Super Bubble	, Jawbusters
```{r}
head(candy[order(candy$winpercent),], n=5)
```

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
head(candy[order(-candy$winpercent),], n=5)

```
> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)
ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```
>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

> Q17. What is the worst ranked chocolate candy?

-The worst ranked chocolate candy is Sixlets


>Q18. What is the best ranked fruity candy?

-The best ranked fruity candy is Starbursts 

4. Taking a look at pricepercent
```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 25)
```

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

```{r}
ord <- order(candy$pricepercent, decreasing = FALSE)
head( candy[ord,c(11,12)], n=5 )

```

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

5 Exploring the correlation structure
```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)

```
>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

-Fruity and chocolate are the most anti correlated
>Q23. Similarly, what two variables are most positively correlated?

-Chocolate and bar are the most positively correlated

6. Principal Component Analysis

```{r}
pca <-prcomp(candy, scale=TRUE)
  summary(pca)

```

```{r}
plot(pca$x[,1:2])
```

```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
#library(plotly)
#ggplotly(p)
```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

-Orignal variables picked up strognly be PC1 in the postive direction is fruity candies and pluribus, and this does make sense since fruit candies are often packaged with multiple other candy pieces, it's just not typically one by itself.