---
title: "Class 5 Data Visual with ggplot2"
author: "Destiny (A16340362)"
format: pdf
---

##Using ggplot

The ggplot2 package needs to be installed as it doesn't come with R "out of the box."

We use the `install.packages()` function to do this.

```{r}
head(cars)
```

To use ggplot, I need to load it up before I can call any of the functions in the package. I do this with the `library()` function.

```{r}
library(ggplot2)
ggplot()
```

ALL ggplot figures have at least 3 things: 
-data(the stuff we want to plot)
-aesthetic mapping (aes values)
-geometry (geom)

```{r}
ggplot(cars)+
  aes(x=speed, y=dist) +
  geom_point()
```

ggplot is not the only graphing system in R, there are a lot of others. There is even "base R" graphics 

```{r}
plot(cars)
```

ggplot(cars)
ggplot(cars)+
aes(x=speed, y=dist)
```{r}
ggplot(cars)+ 
  aes(x=speed, y=dist) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE) +
  labs(title ="Speed and Distances of Cars", caption= "caption", subtitle = "subtitle", x="speed(mph)", y="distance(feet)") +
  theme_bw()
```
```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
nrow(genes)
colnames(genes)
ncol(genes)
table(genes$State)
round(table(genes$State)/nrow(genes)*100, 2)
ggplot(genes) + 
    aes(x=Condition1, y=Condition2) +
    geom_point()
p <- ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point()
p
p + scale_colour_manual( values=c("blue","gray","red") )+labs(y="Drug Treatment", x="Control(no drug")
```

```{r}
library(dplyr)
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"
gapminder <- read.delim(url)
gapminder_2007 <- gapminder %>% filter(year==2007)
gapminder_2007
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha=.5)
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=continent, size=pop) +
  geom_point(alpha=0.5)
ggplot(gapminder_2007) +
  aes(x=gdpPercap, y=lifeExp, color=pop) +
  geom_point(alpha=0.8)
ggplot(gapminder_2007) + 
  aes(x = gdpPercap, y = lifeExp, size = pop) +
  geom_point(alpha=0.5)+ 
  scale_size_area(max_size = 10)
```

```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)
gapminder_1957
ggplot(gapminder_1957) +
  aes(x=gdpPercap, y=lifeExp, color=continent,
                 size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 15)

```
```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)
ggplot(gapminder_1957) +
  aes(x=gdpPercap, y=lifeExp, color=continent,
                 size = pop) +
  geom_point(alpha=0.7) + 
  scale_size_area(max_size = 10) +
  facet_wrap(~year)
```
```{r}
ggplot(mtcars) + aes(x=mpg, y=disp) + geom_point()
```
```{r}
ggplot(mtcars, aes(mpg, disp)) + geom_point()
```

