---
title: "Homework (Class 12 Questions)"
author: "Destiny (A16340362)"
format: pdf
---

>Q13: Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes. 

Sample size for genotype is 462 and the median expression is 25.11561.

Hint: The read.table(), summary() and boxplot() functions will likely be useful here. There is an example R script online to be used ONLY if you are struggling in vein. Note that you can find the medium value from saving the output of the boxplot() function to an R object and examining this object.

There is also the medium() and summary() function that you can use to check your
understanding.

```{r}
expression <- read.table("ExpressionGenotypeResult.txt")
head(expression)
```

```{r}
summary(expression)
```



```{r}
table(expression$geno)
```


```{r}
box <- boxplot(expression$exp)
median(box$stats[3])
```



>Q14: Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3? 

Hint: An example boxplot is provided overleaf – yours does not need
to be as polished as this one.

The SNP does effpect the expression of ORMDL3, since you can see with G/G there is a more reduced expression compared to having A/A.

```{r}
library(ggplot2)
ggplot(expression) +aes(x=geno, y= exp)+ geom_boxplot()
```


